---
output:
  rmarkdown::github_document:
    html_preview: no
---

```{r, include = FALSE}
devtools::load_all()
knitr::opts_chunk$set(fig.path = "man/figures/README-", fig.align = "center")
```

<!--- README.md is generated from README.Rmd. Please edit that file -->

# prioritizrdata: Conservation Planning Datasets

[![lifecycle](https://img.shields.io/badge/Lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![R-CMD-check-Ubuntu](https://img.shields.io/github/workflow/status/prioritizr/prioritizrdata/Ubuntu/master.svg?label=Ubuntu)](https://github.com/prioritizr/prioritizrdata/actions)
[![R-CMD-check-Windows](https://img.shields.io/github/workflow/status/prioritizr/prioritizrdata/Windows/master.svg?label=Windows)](https://github.com/prioritizr/prioritizrdata/actions)
[![R-CMD-check-Mac-OSX](https://img.shields.io/github/workflow/status/prioritizr/prioritizrdata/Mac%20OSX/master.svg?label=Mac%20OSX)](https://github.com/prioritizr/prioritizrdata/actions)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/prioritizrdata)](https://CRAN.R-project.org/package=prioritizrdata)

The _prioritizrdata R_ package is a supplemental package that contains example datasets for conservation planning. It is intended to be used alongside the [_prioritizr R_ package](https://prioritizr.net)---a package for building and solving systematic conservation prioritization problems using integer linear programming (ILP) techniques---and provides little functionality itself.

## Overview

This package contains the following conservation planning datasets.

* __Tasmania dataset__: This dataset was included to show how prioritizations can be created using vector (shapefile) data. It was obtained from the ["Introduction to Marxan" course](https://marxansolutions.org/) and was originally a subset of a larger spatial prioritization project performed under contract to Australia's Department of Environment and Water Resources. It contains vector-based planning unit data and the distributions of 63 vegetation classes in Tasmania, Australia. Refer to the [Tasmania vignette](https://prioritizr.net/articles/tasmania.html) in the _prioritizr R_ package for a worked example with this dataset.
* __Salt Spring Island dataset__: This dataset was included to provide an example of how prioritizations can be created using raster data. It was obtained as part of an online [Marxan-based planning tool](https://arcese.forestry.ubc.ca/marxan-tool/) created for the Coastal Douglas-fir Conservation Partnership (CDFCP). It contains a raster-based planning unit layer and data for five key ecological communities in Salt Spring Island. Refer to the [Salt Spring Island vignette](https://prioritizr.net/articles/saltspring.html) in the _prioritizr R_ package for a worked example with this dataset.

## Installation

The latest official version of the _prioritizrdata R_ package can be installed using the following _R_ code.

```{r, eval = FALSE}
install.packages("prioritizrdata", repos = "https://cran.rstudio.com/")
```

Alternatively, the latest development version can be installed using the following code. Please note that while developmental versions may contain additional features not present in the official version, they may also contain coding errors.

```{r, eval = FALSE}
if (!require(devtools))
  install.packages("devtools")
devtools::install_github("prioritizr/prioritizrdata")
```

```{r, include = FALSE}
# define helpers
is_rel_version <- function(x) {
  length(strsplit(as.character(x), ".", fixed = TRUE)[[1]]) == 3
}

package_version_year <- function(x) {
  if (x %in% rownames(available.packages())) {
    return(live_package_version_year(x))
  } else {
    return(archive_package_version_year(x))
  }
}

live_package_version_year <- function(x) {
  # fetch data
  d <- versions::available.versions(x)[[1]]
  d <- d[which(d$version == max(d$version)), , drop = FALSE]
  c(as.character(d$version), as.character(format(as.POSIXct(d$date), "%Y")))
}

archive_package_version_year <- function(x) {
  # fetch data (copied from remotes R package)
  tf <- tempfile(fileext = ".gz")
  on.exit(unlink(tf), add = TRUE)
  download.file(tf, url = sprintf("%s/src/contrib/Meta/archive.rds", repo))
  con <- gzfile(tf, "rb")
  on.exit(close(con), add = TRUE)
  d <- readRDS(con)[[x]]
  # extract version
  rel_version <- basename(rownames(d))
  rel_version <- sapply(strsplit(rel_version, "_", fixed = TRUE), `[[`, 2)
  rel_version <- sapply(strsplit(rel_version, ".tar", fixed = TRUE), `[[`, 1)
  rel_version <- max(rel_version)
  # extract year
  rel_year <- as.character(max(format(d[, "mtime"], "%Y")))
  # return year version
  c(rel_version, rel_year)
}
```

## Citation

Please cite the _prioritizrdata R_ package when using it in publications. To cite the latest official version, please use:

```{r, include = FALSE}
dev_version_year <- c(
  as.character(packageDescription("prioritizrdata")$Version),
  format(Sys.time(), "%Y")
)
if (is_rel_version(dev_version_year[[1]])) {
  rel_version_year <- dev_version_year
} else {
  rel_version_year <- package_version_year("prioritizrdata")
}
```

```{r, echo = FALSE, results = "asis", comment = ""}
cat(paste0("> Hanson JO, Schuster R, Morrell N, Strimas-Mackey M, Watts ME, Arcese P, Bennett J, Possingham HP (",rel_version_year[[2]],"). prioritizrdata: Conservation Planning Datasets. R package version ", rel_version_year[[1]],". Available at https://CRAN.R-project.org/package=prioritizrdata.\n"))
```

Alternatively, to cite the latest development version, please use:

```{r, echo = FALSE, results = "asis", comment = ""}
cat(paste0("> Hanson JO, Schuster R, Morrell N, Strimas-Mackey M, Watts ME, Arcese P, Bennett J, Possingham HP (", dev_version_year[[2]],"). prioritizrdata: Conservation Planning Datasets. R package version ", dev_version_year[[1]], ". Available at https://github.com/prioritizr/prioritizrdata.\n"))
```
