---
output:
  rmarkdown::github_document:
    html_preview: no
---

```{r, include = FALSE}
devtools::load_all()
knitr::opts_chunk$set(fig.path = "man/figures/README-", fig.align = "center")
```

<!--- README.md is generated from README.Rmd. Please edit that file -->

# prioritizrdata: Conservation Planning Datasets

[![lifecycle](https://img.shields.io/badge/Lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![R-CMD-check-Ubuntu](https://img.shields.io/github/actions/workflow/status/prioritizr/prioritizrdata/R-CMD-check-ubuntu.yaml?branch=main&label=Ubuntu)](https://github.com/prioritizr/prioritizrdata/actions)
[![R-CMD-check-Windows](https://img.shields.io/github/actions/workflow/status/prioritizr/prioritizrdata/R-CMD-check-windows.yaml?branch=main&label=Windows)](https://github.com/prioritizr/prioritizrdata/actions)
[![R-CMD-check-macOS](https://img.shields.io/github/actions/workflow/status/prioritizr/prioritizrdata/R-CMD-check-macos.yaml?branch=main&label=macOS)](https://github.com/prioritizr/prioritizrdata/actions)
[![Documentation](https://img.shields.io/github/actions/workflow/status/prioritizr/prioritizrdata/documentation.yaml?branch=main&label=Documentation)](https://github.com/prioritizr/prioritizrdata/actions)
[![Coverage Status](https://img.shields.io/codecov/c/github/prioritizr/prioritizrdata?label=Coverage)](https://app.codecov.io/gh/prioritizr/prioritizrdata/branch/main)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/prioritizrdata)](https://CRAN.R-project.org/package=prioritizrdata)

The _prioritizrdata R_ package is a supplemental package that contains example datasets for conservation planning. It is intended to be used alongside the [_prioritizr R_ package](https://prioritizr.net).

## Overview

This package contains the following datasets.

* **Tasmania dataset**: Conservation planning dataset for Tasmania, Australia.
* **Salt Spring Island dataset**: Conservation planning dataset for Tasmania, Australia.
* **Washington dataset**: Conservation planning dataset for Washington, The United States of America.

## Installation

The latest official version of the _prioritizrdata R_ package can be installed using the following _R_ code.

```{r, eval = FALSE}
install.packages("prioritizrdata", repos = "https://cran.rstudio.com/")
```

Alternatively, the latest development version can be installed using the following code. Please note that while developmental versions may contain additional features not present in the official version, they may also contain coding errors.

```{r, eval = FALSE}
if (!require(devtools))
  install.packages("devtools")
devtools::install_github("prioritizr/prioritizrdata")
```

## Citation

Please cite the _prioritizrdata R_ package when using it in publications. To cite the latest official version, please use:

```{r, include = FALSE}
# define helpers
is_rel_version <- function(x) {
  length(strsplit(as.character(x), ".", fixed = TRUE)[[1]]) == 3
}

package_version_year <- function(x) {
  if (x %in% rownames(available.packages())) {
    return(live_package_version_year(x))
  } else {
    return(archive_package_version_year(x))
  }
}

live_package_version_year <- function(x) {
  # fetch data
  d <- versions::available.versions(x)[[1]]
  d <- d[which(d$version == max(d$version)), , drop = FALSE]
  c(as.character(d$version), as.character(format(as.POSIXct(d$date), "%Y")))
}

archive_package_version_year <- function(x) {
  # fetch data (copied from remotes R package)
  tf <- tempfile(fileext = ".gz")
  on.exit(unlink(tf), add = TRUE)
  download.file(tf, url = sprintf("%s/src/contrib/Meta/archive.rds", repo))
  con <- gzfile(tf, "rb")
  on.exit(close(con), add = TRUE)
  d <- readRDS(con)[[x]]
  # extract version
  rel_version <- basename(rownames(d))
  rel_version <- sapply(strsplit(rel_version, "_", fixed = TRUE), `[[`, 2)
  rel_version <- sapply(strsplit(rel_version, ".tar", fixed = TRUE), `[[`, 1)
  rel_version <- max(rel_version)
  # extract year
  rel_year <- as.character(max(format(d[, "mtime"], "%Y")))
  # return year version
  c(rel_version, rel_year)
}

dev_version_year <- c(
  as.character(packageDescription("prioritizrdata")$Version),
  format(Sys.time(), "%Y")
)
if (is_rel_version(dev_version_year[[1]])) {
  rel_version_year <- dev_version_year
} else {
  rel_version_year <- package_version_year("prioritizrdata")
}
```

```{r, echo = FALSE, results = "asis", comment = ""}
cat(paste0("> Hanson JO, Schuster R, Morrell N, Strimas-Mackey M, Watts ME, Arcese P, Bennett J, Possingham HP (", rel_version_year[[2]] ,"). prioritizrdata: Conservation Planning Datasets. R package version ", rel_version_year[[1]],". Available at https://CRAN.R-project.org/package=prioritizrdata.\n"))
```

Alternatively, to cite the latest development version, please use:

```{r, echo = FALSE, results = "asis", comment = ""}
cat(paste0("> Hanson JO, Schuster R, Morrell N, Strimas-Mackey M, Watts ME, Arcese P, Bennett J, Possingham HP (", dev_version_year[[2]],"). prioritizrdata: Conservation Planning Datasets. R package version ", dev_version_year[[1]], ". Available at https://github.com/prioritizr/prioritizrdata.\n"))
```
